---
title: "Portfolio 3"
author: "Thomas Huang"
date: "2026-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This piece reflects something fancy I recently learned, creating a word cloud in R. I will use the taylor package as the source of my data. Some text analyses in this piece builds on one of my undergraduate course project.

## Load packages

```{r load-packages}
library(taylor);library(tidyverse);library(tidytext);library(quanteda);library(quanteda.textplots)
```

## Extract data

```{r data}
data(taylor_album_songs)
glimpse(taylor_album_songs)

# The lyrics column is nested. I will unnest it.
taylor_album_song.un<- taylor_album_songs %>% 
    unnest(lyrics) 

glimpse(taylor_album_song.un)
```

## Convert lyrics into Quanteda corpus

```{r preprocess-corpus}
lyrics.cp <- corpus(taylor_album_song.un$lyric,
                  docvars = taylor_album_song.un$track_name)
```

## Tokenize

This step removes the punctuation, removing the stopwords, removing the numbers
```{r preprocess-token}
token_lyrics.clean <- lyrics.cp %>% 
  quanteda::tokens(remove_punct = TRUE, 
                   remove_numbers = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(stopwords("en")) 
```

## Create a DFM

A DFM refers to a A Document-Feature Matrix. It transforms unstructured text into structured data based on the tokens. Details can be found here: http://library.virginia.edu/data/articles/a-beginners-guide-to-text-analysis-with-quanteda

```{r preprocess-DFM}
lyrics_dfm <- dfm(token_lyrics.clean)
```

## Word cloud

Feeding the DFM into the textplot function would create the word cloud.

```{r wordcloud}
textplot_wordcloud(
      lyrics_dfm,
      min_size = 0.5,
      max_size = 5,
      min_count = 10,
      max_words = 500,
      color = "RED",
      font = NULL,
      adjust = 0,
      rotation = 0.1
  )
```

