---
title: "Portfolio 4"
author: "Thomas Huang"
date: "2026-03-31"
output: github_document
---

This portfolio piece is part of my data analysis for my thesis. This piece deals with two things: data cleaning and running two two-way ANOVAs.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages and data

```{r}
library(tidyverse);library(readxl);library(psych);library(emmeans);library(effectsize)

df <- read_excel("data_p4.xlsx")
fail_check <- read_excel("Fail_Check.xlsx")
```

## Cleaning

```{r}
glimpse(df)
```

## Merge scenarios

```{r}
# Recode 2 into 0 in scenario items
df <- df %>%
  mutate(across(paste0("SPR", 1:5),
                ~ ifelse(. == 2, 0, .))) %>% 
  mutate(across(paste0("SPB", 1:5),
                ~ ifelse(. == 2, 0, .))) %>% 
  mutate(across(paste0("OPR", 1:5),
                ~ ifelse(. == 2, 0, .))) %>% 
  mutate(across(paste0("OPB", 1:5),
                ~ ifelse(. == 2, 0, .)))

df <- df %>%
  mutate(
    S1 = rowSums(select(., SPR1, SPB1, OPR1, OPB1), na.rm = TRUE),
    S2 = rowSums(select(., SPR2, SPB2, OPR2, OPB2), na.rm = TRUE),
    S3 = rowSums(select(., SPR3, SPB3, OPR3, OPB3), na.rm = TRUE),
    S4 = rowSums(select(., SPR4, SPB4, OPR4, OPB4), na.rm = TRUE),
    S5 = rowSums(select(., SPR5, SPB5, OPR5, OPB5), na.rm = TRUE)
  )

df %>% 
  select(S1:S5) %>% 
  psych::describe()
```

## Calculate risk-taking index

```{r}
df$rtidx <- df %>% 
  select(S1:S5) %>% 
  rowMeans(na.rm = TRUE)
```

## Merge other variables

```{r}
# all variable names
vars <- names(df)

# keep only condition-prefixed variables
conds <- c("SPR", "SPB", "OPR", "OPB")

pattern <- paste0("^(", paste(conds, collapse="|"), ")_")

# extract suffix (item names)
items <- unique(sub(pattern, "", vars[grepl(pattern, vars)]))

library(dplyr)

process_all <- function(df, conds) {
  
  pattern <- paste0("^(", paste(conds, collapse="|"), ")_")
  vars <- names(df)
  
  items <- unique(sub(pattern, "", vars[grepl(pattern, vars)]))
  
  for (item in items) {
    
    cols <- paste0(conds, "_", item)
    cols <- cols[cols %in% names(df)]  # only keep existing ones
    
    # skip if less than 2 columns (safety)
    if (length(cols) == 0) next
    
    # v0
    df[[paste0(item, "_v0")]] <- coalesce(!!!df[cols])
    
    # C
    df[[paste0(item, "_C")]] <- conds[
      max.col(!is.na(df[cols]), ties.method = "first")
    ]
  }
  
  return(df)
}

df <- process_all(df, c("SPR", "SPB", "OPR", "OPB"))
df <- df %>% 
  mutate(Condition = MC1_C) %>% 
  select(-ends_with("_C")) %>% 
  rename(
    MC1 = MC1_v0,
    MC2 = MC2_v0,
    AC1 = AC1_v0,
    IM  = IM_v0,
    AC2 = AC2_v0
  )
```

## Define conditions

```{r}
df$target <- ifelse(df$Condition == "OPR" | df$Condition == "OPB", 1, 0) # 0 = self, 1 = other
df$visibility <- ifelse(df$Condition == "SPB" | df$Condition == "OPB", 1, 0) # 0 = private, 1 = public
``` 

## Attention checks

```{r}
failed <- fail_check %>% 
  dplyr::pull(PID) %>% 
  as.list()

df <- df %>% 
  filter(AC2 == 3) %>% 
  filter(PROLIFIC_PID %in% failed == FALSE) # N = 995
```

## Inclusion criteria

```{r}
df <- df %>% 
  filter(Age >= 18 & Age <= 35) %>% 
  filter(Gender == 2) %>% 
  filter(SexualOrientation == 1 | SexualOrientation == 2) # N = 953
```

## Manipulation checks

```{r}
# Self-private
table(df$SPR_MC1)
table(df$SPR_MC2)

# Self-public
table(df$SPB_MC1)
table(df$SPB_MC2)

# Other-private
table(df$OPR_MC1)
table(df$OPR_MC2)

# Other-public
table(df$OPB_MC1)
table(df$OPB_MC2)

fail_MC <- df %>% 
  filter(
    MC1 == 2 & visibility == 0 | 
      MC2 == 2 & visibility == 0 |
      MC1 == 1 & visibility == 1 |
      MC2 == 1 & visibility == 1)

table(fail_MC$Condition)

df2 <- df
save(df2, file = "df2.RData")

df <- df %>% 
  filter(MC1 == 1 & visibility == 0 | MC1 == 2 & visibility == 1) %>% 
  filter(MC2 == 1 & visibility == 0 | MC2 == 2 & visibility == 1) #N = 922. 31 participants did not pass MC.
```

## Risk-taking across conditions

```{r}
df$Condition <- factor(df$Condition)
df$target <- factor(df$target)
df$visibility <- factor(df$visibility)

D_rt <- psych::describeBy(df$rtidx, df$Condition)
D_rt

psych::describeBy(df$rtidx, df$target)
psych::describeBy(df$rtidx, df$visibility)

model_rt <- aov(rtidx ~ target * visibility, data = df)
summary(model_rt)

# Simple effects within decision target
joint_tests(model_rt, by = "target")

# Simple effects within visibility
joint_tests(model_rt, by = "visibility")

# Effect sizes
eta_squared(model_rt, partial = TRUE)

# Effect size of other-private vs other-public
L <- D_rt[["OPB"]][["mean"]] - D_rt[["OPR"]][["mean"]]
SS_effect <- L^2/(1/D_rt[["OPB"]][["n"]] + 1/D_rt[["OPR"]][["n"]])
aov_tab <- anova(model_rt)
SS_error  <- aov_tab["Residuals", "Sum Sq"]
rt_eta2_p <- SS_effect/(SS_effect+SS_error)
rt_eta2_p

# Effect size of self-private vs self-public
L <- D_rt[["SPB"]][["mean"]] - D_rt[["SPR"]][["mean"]]
SS_effect <- L^2/(1/D_rt[["SPB"]][["n"]] + 1/D_rt[["SPR"]][["n"]])
aov_tab <- anova(model_rt)
SS_error  <- aov_tab["Residuals", "Sum Sq"]
rt_eta2_p <- SS_effect/(SS_effect+SS_error)
rt_eta2_p

# Effect size of self-private vs other-private
L <- D_rt[["OPR"]][["mean"]] - D_rt[["SPR"]][["mean"]]
SS_effect <- L^2/(1/D_rt[["OPR"]][["n"]] + 1/D_rt[["SPR"]][["n"]])
aov_tab <- anova(model_rt)
SS_error  <- aov_tab["Residuals", "Sum Sq"]
rt_eta2_p <- SS_effect/(SS_effect+SS_error)
rt_eta2_p

# Effect size of self-public vs other-public
L <- D_rt[["OPB"]][["mean"]] - D_rt[["SPB"]][["mean"]]
SS_effect <- L^2/(1/D_rt[["OPB"]][["n"]] + 1/D_rt[["SPB"]][["n"]])
aov_tab <- anova(model_rt)
SS_error  <- aov_tab["Residuals", "Sum Sq"]
rt_eta2_p <- SS_effect/(SS_effect+SS_error)
rt_eta2_p
```

## Perceived impression management concerns across conditions

```{r}
# Compare impression management
D_im <- psych::describeBy(df$IM, df$Condition)
D_im

psych::describeBy(df$IM, df$target)
psych::describeBy(df$IM, df$visibility)

model_im <- aov(IM ~ target * visibility, data = df)
summary(model_im)

# Simple effects within decision target
joint_tests(model_im, by = "target")

# Simple effects within visibility
joint_tests(model_im, by = "visibility")

# Effect sizes
eta_squared(model_im, partial = TRUE)

# Effect size of other-private vs other-public
L <- D_im[["OPB"]][["mean"]] - D_im[["OPR"]][["mean"]]
SS_effect <- L^2/(1/D_im[["OPB"]][["n"]] + 1/D_im[["OPR"]][["n"]])
aov_tab <- anova(model_im)
SS_error  <- aov_tab["Residuals", "Sum Sq"]
im_eta2_p <- SS_effect/(SS_effect+SS_error)
im_eta2_p

# Effect size of self-private vs self-public
L <- D_im[["SPB"]][["mean"]] - D_im[["SPR"]][["mean"]]
SS_effect <- L^2/(1/D_im[["SPB"]][["n"]] + 1/D_im[["SPR"]][["n"]])
aov_tab <- anova(model_im)
SS_error  <- aov_tab["Residuals", "Sum Sq"]
im_eta2_p <- SS_effect/(SS_effect+SS_error)
im_eta2_p

# Effect size of self-private vs other-private
L <- D_im[["OPR"]][["mean"]] - D_im[["SPR"]][["mean"]]
SS_effect <- L^2/(1/D_im[["OPR"]][["n"]] + 1/D_im[["SPR"]][["n"]])
aov_tab <- anova(model_im)
SS_error  <- aov_tab["Residuals", "Sum Sq"]
im_eta2_p <- SS_effect/(SS_effect+SS_error)
im_eta2_p

# Effect size of self-public vs other-public
L <- D_im[["OPB"]][["mean"]] - D_im[["SPB"]][["mean"]]
SS_effect <- L^2/(1/D_im[["OPB"]][["n"]] + 1/D_im[["SPB"]][["n"]])
aov_tab <- anova(model_im)
SS_error  <- aov_tab["Residuals", "Sum Sq"]
im_eta2_p <- SS_effect/(SS_effect+SS_error)
im_eta2_p
```

## Compare risk-taking by scenario

```{r}
psych::describeBy(df$S1, df$Condition) #Dance
psych::describeBy(df$S2, df$Condition) #Talk to a guy
psych::describeBy(df$S3, df$Condition) #Date a guy at gym
psych::describeBy(df$S4, df$Condition) #Valentine's Day gift
psych::describeBy(df$S5, df$Condition) #Confess to a close friend
```

## Save data for mediation analysis

```{r}
save(df, file = "df.RData")
```
