---
title: "p07"
author: "Thomas Huang"
date: "2026-05-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This portflio piece does different kinds of automated bivariate association tests for an exploratory project. Being automated means there is no need to running the tests one by one. Instead, it can return the test results with Bonferroni correction and export the results in Excel sheets just by inputting the IVs and DV. 

This piece runs Pearson's r (continuous IV and DV), ANOVA (categorical IV and continuous DV), and chi-square test (categorical IV and DV).

# Load packages and data

```{r}
library(tidyverse);library(vcd);library(gmodels);library(openxlsx);library(Hmisc);library(car);library(MOTE)
load("merged_650_r.RData")
```

# Pearson's r

Pearson's r is easy in this case. There is a ready-made function that takes a data frame and returns a correlation matrix.

## Subset data

```{r}
x <- merged_650_r

x <- x %>% 
  dplyr::select(mCO1_ImpactOnRelationship, mCO1_ImpactOnLGBTQ, # Two DVs
                mAge, mSex, mEthnicMinority, mOnlyChild, mSingle:mRelationshipStatus_Other, mAttitudeByProvince, mSocialLadder,
                mCO1_OtherGender, mCO1_Age, mCO1_Medium_1:mCO1_Medium_888, mCO1_Context_1:mCO1_Context_888, mCO1_Closeness,
                mCO1_Strategy_1:mCO1_Strategy_5_31) # IVs
```

## Calculate correlation matrix

```{r}
x <- as.data.frame(lapply(x, function(col) as.numeric(haven::zap_labels(col))))
numeric <- data.matrix(x)
corrmatrix_pearson <- Hmisc::rcorr(numeric)
```

## View correlation coefficient

```{r}
r <- round(corrmatrix_pearson$r,2) %>% 
  as.data.frame 
r <- select(r, mCO1_ImpactOnRelationship, mCO1_ImpactOnLGBTQ) # Select the two columns that are the two continuous outcomes.
r <- r[-c(1, 2), ] # Drop the first and second rows that are the two outcomes
r
```

## View p-value

```{r}
p <- corrmatrix_pearson$P %>%
  as.data.frame 
p <- select(p, mCO1_ImpactOnRelationship, mCO1_ImpactOnLGBTQ) # Select the two columns that are the two continuous outcomes.
p <- p[-c(1, 2), ] # Drop the first and second rows that are the two outcomes.

# Conduct Bonferroni correction.
p_adj <- mutate_all(p, function(x) x * (3*85)) # There are 3*85 comparisons (Three outcomes and 85 predictors).

# Convert p-values larger than 1 into ">.999"
p$mCO1_ImpactOnRelationship <- ifelse(p$mCO1_ImpactOnRelationship >.999, 
                                            ">.999", 
                                          ifelse(p$mCO1_ImpactOnRelationship <.001, 
                                                    "<.001", 
                                                    round(p$mCO1_ImpactOnRelationship, 3)
                                              )
                                          )

p$mCO1_ImpactOnLGBTQ <- ifelse(p$mCO1_ImpactOnLGBTQ >.999, 
                              ">.999", 
                              ifelse(p$mCO1_ImpactOnLGBTQ <.001, 
                                      "<.001", 
                                      round(p$mCO1_ImpactOnLGBTQ, 3)
                                )
  )

p_adj$mCO1_ImpactOnRelationship <- ifelse(p_adj$mCO1_ImpactOnRelationship >.999, 
                                            ">.999", 
                                          ifelse(p_adj$mCO1_ImpactOnRelationship <.001, 
                                                    "<.001", 
                                                    round(p_adj$mCO1_ImpactOnRelationship, 3)
                                              )
                                          )

p_adj$mCO1_ImpactOnLGBTQ <- ifelse(p_adj$mCO1_ImpactOnLGBTQ >.999, 
                              ">.999", 
                              ifelse(p_adj$mCO1_ImpactOnLGBTQ <.001, 
                                      "<.001", 
                                      round(p_adj$mCO1_ImpactOnLGBTQ, 3)
                                )
  )
p

p_adj #Bonferroni-corrected p
```

## View sample sizes after listwise deletion

```{r}
n <- round(corrmatrix_pearson$n,1) %>% 
  as.data.frame
n <- select(n, mCO1_ImpactOnRelationship, mCO1_ImpactOnLGBTQ) # Select the two columns that are the two continuous outcomes.
n <- n[-c(1, 2), ] # Drop the first and second rows that are the two outcomes
n
```

## Create an Excel sheet

```{r}
Pearson_r.table <- data.frame(
  variables = rownames(r),
  r_mCO1_ImpactOnRelationship = r$mCO1_ImpactOnRelationship,
  p_mCO1_ImpactOnRelationship = p$mCO1_ImpactOnRelationship,
  padj_mCO1_ImpactOnRelationship = p_adj$mCO1_ImpactOnRelationship,
  n_mCO1_ImpactOnRelationship = n$mCO1_ImpactOnRelationship,
  r_mCO1_ImpactOnLGBTQ = r$mCO1_ImpactOnLGBTQ,
  p_mCO1_ImpactOnLGBTQ = p$mCO1_ImpactOnLGBTQ,
  padj_mCO1_ImpactOnLGBTQ = p_adj$mCO1_ImpactOnLGBTQ,
  n_mCO1_ImpactOnLGBTQ = n$mCO1_ImpactOnLGBTQ
)

# Save the table S7 to excel sheet
write.xlsx(Pearson_r.table, "Pearson_r.table.xlsx")
```

# ANOVA

There is no ready-made automation functions. So I will define a function and a loop. The function will return test statistics, two df's, effect size, p value, and Bonferroni corrected p values.

## Subset data

```{r}
df_anova <- merged_650_r
df_anova <- df_anova %>% 
  dplyr::select(
    mCO1_ImpactOnRelationship, mCO1_ImpactOnLGBTQ,
    mDegree, mEmployment, mSocialAssistance,
    mCO1_SexualOrientation, mCO1_Relationship, mCO1_Number_Perceivers, mCO1_Occasion, mCO1_Venue, mCO1_Dialogue, 
    mCO1_Suspect, mCO1_Imagine, mCO1_Hope
  ) %>%
  as_factor() %>% 
  mutate(
    mCO1_ImpactOnRelationship = as.numeric(mCO1_ImpactOnRelationship)
  )

glimpse(df_anova)

# Create a df with all IVs but without the DV
df_anova_r <- df_anova %>% 
  dplyr::select(-mCO1_ImpactOnRelationship, -mCO1_ImpactOnLGBTQ)
```

## Define function and loop

```{r results='hide'}
ANOVA.table <- data.frame(
  variable = character(0),
  df1 = numeric(0),
  df2 = numeric(0),
  f = numeric(0),
  p = numeric(0),
  adjusted_p = numeric(0),
  omega2 = numeric(0)
)

for (var_i in colnames(df_anova_r)){
  print(paste("===============================", var_i, "==============================="))
  df <- df_anova
  x <- df[[var_i]]
  y <- df_anova$mCO1_ImpactOnRelationship

  # Perform ANOVA
  model <- lm(y ~ x, data=df)  
  anova_result <- Anova(model, type="III")
  
  # Print degrees of freedom
  df1 <- anova_result$`Df`[2]
  df1
  df2 <- anova_result$`Df`[3]
  df2
  
  # Print F-value
  f <- round(anova_result$`F value`[2], 2)
  f
  
  # Print p-value
  p <- ifelse(anova_result$`Pr(>F)`[2] > .999, 
                                ">.999", 
                                ifelse(anova_result$`Pr(>F)`[2] <.001, 
                                       "<.001", 
                                        round(anova_result$`Pr(>F)`[2], 3)
                                 )
   )
  p
  
  # Perform Bonferroni correction
  adjusted_p <- anova_result$`Pr(>F)`[2] * (3 * 85)
  adjusted_p <- ifelse(adjusted_p > .999, 
                                ">.999", 
                                ifelse(adjusted_p <.001, 
                                        "<.001", 
                                        round(adjusted_p, 3)
                                 )
  )
  adjusted_p
  
  # Calculate omega squared
  model <- aov(y ~ x, data=df)  
  SS_b <- summary(model)[[1]]$`Sum Sq`[1]  # Sum of squares between groups
  df_b <- summary(model)[[1]]$Df[1]       # Degrees of freedom between groups
  SS_w <- summary(model)[[1]]$`Sum Sq`[2]  # Sum of squares within groups
  df_w <- summary(model)[[1]]$Df[2]       # Degrees of freedom within groups
  MS_w <- SS_w / df_w                     # Mean square within groups
  omega2 <- (SS_b - df_b * MS_w) / (SS_b + SS_w + MS_w)
  omega2 <- round(omega2, 3)

  # Combine the row (two dfs and some values) to a blank df
  ANOVA.table_row <- cbind(var_i, df1, df2, f, p, adjusted_p, omega2)
  
  names(ANOVA.table_row) <- names(ANOVA.table)

  ANOVA.table <- rbind(ANOVA.table, ANOVA.table_row)
  glimpse(ANOVA.table) 
}
```

## Create an Excel sheet

```{r}
ANOVA.table

# Save the table S7 to excel sheet
write.xlsx(ANOVA.table, "ANOVA.table.xlsx")
```

# Chi-square tests

Similar to the ANOVAs, I defined a function and a loop that do the magic.

## Subset data

```{r}
df_chisq <- merged_650_r %>% 
  dplyr::select(
                mSex, mEthnicMinority, mOnlyChild, mDegree, mSingle:mWidowed, mRelationshipStatus_Other,
                mEmployment, mSocialAssistance, 
                mCO1_OtherGender:mCO1_Strategy_5_31
  ) %>% 
  dplyr::select(-mWidowed,
                -mCO1_SexualOrientation_TEXT, -mCO1_Context_Other_TEXT, -mCO1_Reaction_Other_TEXT,
                -mCO1_Age,-(mCO1_Reaction_1:mCO1_ImpactOnLGBTQ)
  )
glimpse(df_chisq)
```

## Define a function and a loop

```{r results='hide'}
chi.table <- data.frame(
  variable = character(0),
  chisq = numeric(0),
  p = numeric(0),
  adjusted_p = numeric(0),
  cramer = numeric(0)
)

for (var_i in colnames(df_chisq)){
  print(paste("===============================", var_i, "==============================="))
  df <- df_chisq
  x <- df[[var_i]]
  y <- merged_650_r$mCO1_Reaction_LCA_class

  # Run Chi-Sq analysis
  # Use tryCatch to handle errors
result <- tryCatch(
  {
    CrossTable(table(x,y),expected=T,prop.r=F,prop.c=T,prop.t=F,prop.chisq=F,chisq=T,fisher=T,format="SPSS")
  },
  error = function(e) {
    # If an error occurs
    CrossTable(table(x,y),expected=T,prop.r=F,prop.c=T,prop.t=F,prop.chisq=F,chisq=T,simulate.p.value=TRUE, format="SPSS")
  }
)
  

  # Create a contingency table
  chi_sq_result <- table(x, y) %>%
    chisq.test()

  p_value <- ifelse(chi_sq_result$p.value > .999, 
                              ">.999", 
                              ifelse(chi_sq_result$p.value <.001, 
                                      "<.001", 
                                      round(chi_sq_result$p.value, 3)
                                )
  )

  # Conduct Bonferroni correction
  adjusted_p_value <- chi_sq_result$p.value * (3 * 85)
  adjusted_p_value <- ifelse(adjusted_p_value > .999, 
                              ">.999", 
                              ifelse(adjusted_p_value <.001, 
                                      "<.001", 
                                      round(adjusted_p_value, 3)
                                )
  )

  # Print the adjusted p-value
  print(adjusted_p_value)

  # Calculate Cramer's V
  tab1 <- gmodels::CrossTable(x,y)
  chisq_summary <- vcd::assocstats(tab1$t)
  chisq_value <- round(chisq_summary$chisq_tests["Pearson", "X^2"], 2)
  cramer_value <- round(chisq_summary$cramer, 2)

    # Combind the row (two dfs and some values) to a blank df
  chi.table_row <- cbind(var_i, chisq_value, p_value, adjusted_p_value, cramer_value)
  
  names(chi.table_row) <- names(chi.table)

  chi.table <- rbind(chi.table, chi.table_row)
  glimpse(chi.table) 
}
```

## Create an Excel sheet

```{r}
chi.table

# Save the table S7 to excel sheet
write.xlsx(chi.table, "chi.table.xlsx")
```