---
title: "p10"
author: "Thomas Huang"
date: "2026-05-05"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This portfilio piece aims to investigate a classic topic in marketing: whether people think neagtive or positive are more useful. To answer this question, I used a slice of Yelp's open dataset, US restaurant reviews in July 2019. I did not use the full dataset. Otherwise, my laptop would freak out.

My MO is to find out whether a review is positive or negativw through text analysis with quanteda and then test whether people marked them as useful.

# Load packages and data

```{r}
library(tidyverse);library(quanteda);library(quanteda.textstats)

load('yelp.cleaned.RData')
glimpse(yelp.cleaned)
```

# Text analysis
## Preprocessing

```{r}
review.cp <- corpus(yelp.cleaned$text,
                  docvars = yelp.cleaned$Code)

summary(review.cp, 5)

# Tokenization
token.clean <- review.cp %>% 
  quanteda::tokens(remove_punct = TRUE, 
                   remove_numbers = TRUE) %>%
  tokens_tolower() %>%
  tokens_remove(stopwords("en")) 

# Create a DFM
review_dfm <- dfm(token.clean)
```

## Sentiment analysis

I'm running sentiment analysis by comparing the tokens with a dictionary that marks each token as positive, neutral, or negative.

```{r}
data("data_dictionary_LSD2015")

sent_score.dfm <- dfm_lookup(review_dfm,
                                    dictionary = data_dictionary_LSD2015) 

sent_score <- convert(sent_score.dfm, to = "data.frame")

yelp.cleaned <- sent_score %>% 
  dplyr::mutate(Code = 1:83286) %>%
  dplyr::select(negative,positive, Code) %>% 
  left_join(yelp.cleaned, by = "Code") %>%
  relocate(negative:positive, .after = last_col())
```

# See how sentiment predicts useful rating

```{r}
model <- lm(useful ~ negative + positive, data = yelp.cleaned)
summary(model)
```

This simple model tells that both positive and negative sentiments positively predicted useful ratings. I would like to add some covariates

# Calculate the length of reviews

```{r}
yelp.cleaned <- textstat_summary(review.cp) %>% 
  dplyr::mutate(Code = 1:83286) %>% 
  dplyr::select(Code, chars) %>%
  left_join(yelp.cleaned, by = "Code") %>%
  relocate(chars, .after = last_col())
```

# Calculate diversity score (TTR)

```{r}
yelp.cleaned <- textstat_lexdiv(
  review_dfm,
  measure = "TTR",
  remove_numbers = TRUE,
  remove_punct = TRUE,
  remove_symbols = TRUE,
  remove_hyphens = FALSE,
  log.base = 10,
  MATTR_window = 100L,
  MSTTR_segment = 100L
) %>% 
  as.data.frame() %>%
  dplyr::mutate(Code = 1:83286) %>% 
  dplyr::select(Code, TTR) %>%
  left_join(yelp.cleaned, by = "Code") %>%
  relocate(TTR, .after = last_col())
```

# Run a more complex model

```{r}
model2 <- lm(useful ~ negative + positive + chars*negative + chars*positive + TTR, data = yelp.cleaned)
summary(model2)
```

Here are some more interesting findings. After controlling for length and diversity score, negative sentiment predicted higher usefulness ratings, but positive sentiment had little effect on usefulness ratings. But if a negative review gets longer, the usefulness rating is predicted to decrease. In contrast, if a positive review gets longer, the usefulness rating is predicted to increase.
