---
title: "Replicating a two-way ANOVA in a regression"
author: "Thomas Huang"
date: "2026-04-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/roberthuang/Desktop/Wake/Spring 2026/DSPsych/dspych_portfolio/p05")
```

The cool thing with this project is that it replicates the exact results of a two-way ANOVA.

## Packages and data

```{r}
library(tidyverse);library(MplusAutomation);library(here)
load("df.RData")

colnames(df)
```

First, I will show how a continuous variable is regressed on two factors. Usually, this is done with dummy coding.

## Rename variables

```{r}
df$t <- ifelse(df$Condition == "OPR" | df$Condition == "OPB", 0, 1) # 0 = self, 1 = other
df$v <- ifelse(df$Condition == "SPB" | df$Condition == "OPB", 0, 1) # 0 = private, 1 = public
df$vt <- df$v*df$t
df$rt <- df$rtidx

df %>% 
  select(v, t) %>% 
  psych::describe()
```

## Regression with dummy coding

In this case, the coefficients of the IVs are simple slopes, or simple effects in an ANOVA. Therefore, what is problematic here is that I cannot estimate the main effects like an ANOVA. A potential consequence is that, if I would like to run a mediation model extending from this regression, I cannot estimate the indirect effects of the main effects of the IVs.
```{r}
reg_dummy <- lm(rtidx ~ v + t + vt, data = df)
summary(reg_dummy)
```

## Replicating the ANOVA with a smarter coding scheme

A way to solve this problem is to ditch dummy coding and to code the IVs as -0.5 and 0.5

```{r}
df$t <- ifelse(df$Condition == "OPR" | df$Condition == "OPB", -0.5, 0.5) # 0 = self, 1 = other
df$v <- ifelse(df$Condition == "SPB" | df$Condition == "OPB", -0.5, 0.5) # 0 = private, 1 = public
df$vt <- df$v*df$t
df$rt <- df$rtidx
```

```{r}
reg_anova <- lm(rtidx ~ v + t + vt, data = df)
summary(reg_anova)
```

## Let's check if this has replicated the ANOVA results

```{r}
D_rt <- psych::describeBy(df$rtidx, df$Condition)
D_rt

anova_rt <- aov(rtidx ~ target * visibility, data = df)
summary(anova_rt)
```

The regression replicated the ANOVA results. Built upon this, we can do two things. First, we can run mediation if we are interested in the indirect effects of the ANOVA main effects. Second, this simply makes ANCOVAs more intuitive. No need to learn how to run an ANCOVA. Just add controls into the above regression.