---
title: "p09"
author: "Thomas Huang"
date: "2026-05-05"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/roberthuang/Desktop/Wake/Spring 2026/DSPsych/dspych_portfolio/p09")
```

In this piece, I'm predicting whether individuals would purchase a discounted premium membership at a supermarket with random forest. This is the soruce of my dataset: https://www.kaggle.com/datasets/ahsan81/superstore-marketing-campaign-dataset/data.

This is the code book:

Response (target) - 1 if customer accepted the offer in the last campaign, 0 otherwise
ID - Unique ID of each customer
Year_Birth - Age of the customer
Complain - 1 if the customer complained in the last 2 years
Dt_Customer - date of customer's enrollment with the company
Education - customer's level of education
Marital - customer's marital status
Kidhome - number of small children in customer's household
Teenhome - number of teenagers in customer's household
Income - customer's yearly household income
MntFishProducts - the amount spent on fish products in the last 2 years
MntMeatProducts - the amount spent on meat products in the last 2 years
MntFruits - the amount spent on fruits products in the last 2 years
MntSweetProducts - amount spent on sweet products in the last 2 years
MntWines - the amount spent on wine products in the last 2 years
MntGoldProds - the amount spent on gold products in the last 2 years
NumDealsPurchases - number of purchases made with discount
NumCatalogPurchases - number of purchases made using catalog (buying goods to be shipped through the mail)
NumStorePurchases - number of purchases made directly in stores
NumWebPurchases - number of purchases made through the company's website
NumWebVisitsMonth - number of visits to company's website in the last month
Recency - number of days since the last purchase

The goal is to find out which are the most important predictors.

# Load packages and data

```{r}
library(tidyverse);library(ggplot2);library(randomForest);library(REAT);library(partykit);library(caret)

store <- read.csv('superstore_data.csv')
```

# Data management

```{r}
glimpse(store)

store$Dt_Customer <- as.Date(store$Dt_Customer, format = "%m/%d/%Y")
store$join_duration <- as.numeric(difftime(Sys.Date(), store$Dt_Customer, units = "days"))
store$join_duration <- store$join_duration/365.25 - 1

store$Age <- 2023 - store$Year_Birth #Data from 2023

store$Education <- as.factor(store$Education)
store$Marital_Status <- as.factor(store$Marital_Status)
store$Response <- as.factor(store$Response)

glimpse(store)
```

# Impute missing values

```{r}
set.seed(4959)
store_imputed <- rfImpute(Response ~
          Age + Education + Marital_Status + Income + Kidhome + Teenhome + join_duration + Recency
          + MntWines + MntFruits + MntMeatProducts + MntFishProducts + MntSweetProducts + MntGoldProds
          + NumDealsPurchases + NumWebPurchases + NumCatalogPurchases + NumStorePurchases + NumWebVisitsMonth + Complain,
          data = store)
```

# Run random forest

```{r}
set.seed(4959)
store_rf <- randomForest(Response ~
          Age + Education + Marital_Status + Income + Kidhome + Teenhome + join_duration + Recency
          + MntWines + MntFruits + MntMeatProducts + MntFishProducts + MntSweetProducts + MntGoldProds
          + NumDealsPurchases + NumWebPurchases + NumCatalogPurchases + NumStorePurchases + NumWebVisitsMonth + Complain,
          ntree = 3000,
          importance = TRUE,
          proximity = TRUE,
          keep.forest = TRUE,
          keep.inbag = TRUE,
          mtry = 50, #These are default settings of the function
          data = store_imputed) 
```

# See if there are enough trees

Looks good.
```{r}
plot(store_rf)
```

# Evaluate model performance

Accuracy = .871. Not bad.
```{r}
# Predicted outcomes
store_predicted <- predict(store_rf, OOB = TRUE, predict.all = TRUE, proximity = TRUE) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Number") %>% 
  as_tibble()

# Confusion matrix
set.seed(4959)
caret::confusionMatrix(store_predicted$pred, store_imputed$Response, positive = '1')
```

# Calculate variable importance

This automatically returns an ugly plot. Please ignore.
```{r}
vimp <- randomForest::varImpPlot(store_rf,
                         sort=FALSE,
                         main="Variable Importance Plot") %>% 
  as.data.frame()

vimp <- vimp %>% 
  as.data.frame %>% 
  rownames_to_column(var = "Variable") %>% 
  as_tibble()

vimp #Both mean decrease in accuracy and mean decrease Gini represent variable importance. I will use the former because decrease in accuracy is used as the loss function in training.
```

```{r}
color_scheme <- ifelse(vimp$MeanDecreaseAccuracy < 0, "pink", "lightblue")

vimp_plot <- ggplot(vimp, aes(x = reorder(Variable, MeanDecreaseAccuracy), y = MeanDecreaseAccuracy)) +
  geom_bar(stat = "identity",
           show.legend = FALSE,
           fill = color_scheme,     
           color = "white",
           width = 1) +
  geom_hline(yintercept = 0, color = 1, lwd = 0.2) +
  geom_text(aes(label = Variable, # Text with groups
                hjust = ifelse(MeanDecreaseAccuracy < 0, 1.2, -0.5),
                vjust = 0.3), size = 4) +
  xlab("Variable") +
  ylab("Mean Decrease in Accuracy") +
  scale_y_continuous(breaks = seq(-10, 100, by = 10),
                     limits = c(-20, 100)) +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        panel.grid.major.y = element_blank())

ggsave("vimp_plot.png", plot = vimp_plot, width = 24, height = 4)
```

Takeaway: Target those who visited recently, spent a lot on wine, and rich.